---
title: "Main Functions"
format: html
---

```{r}
library(Plarrr)
plastics <- load_data()
```

## Planned Functions

### 1. Summarize top plastic-producing companies by country

```{r}
summarize_top_companies(plastics, country = "United States", year = 2020, top_n = 5)
```

Filters to a given country and year, groups by "parent_company", sums
"grand_total", and returns a ranked table of the top offending companies.
Removes filler entries like "Grand Total" and "Unbranded" automatically.

### 2. Plot country-level plastic waste by year

```{r}
plot_country_trend(plastics, countries = c("China", "Philippines", "Nigeria"))
```

Produces a ggplot2 grouped bar chart comparing "grand_total" plastic waste
collected in 2019 vs 2020 for one or more countries side by side. Handles
filtering, grouping, and themes automatically so you just include a character
vector of country names or just a single country name.

### 3. Summarize plastic type breakdown across all countries

```{r}
summarize_plastic_types(plastics, year = 2020)
```

Pivots the seven plastic type columns (hdpe, ldpe, o, pet, pp, ps, pvc) to a
long data frame, sums each type globally for either 2019 or 2020, and
returns a two-column table of plastic type and total count. Automatically
removes filler rows like "Grand Total" and "Unbranded" as well.
